---
title: "sydneys_markdown"
output: html_document
date: "2025-07-14"
---

```{r}
#Authors: [Sydney S]
#Author Date: 07/09/2025
#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 07/14/2025

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(readr)
```

#Loading Datasets
```{r}
#Womens basketball datasets
Incident_B <- read.csv("data-sets/Incident Report WBB.csv")
Dynamo_B <- read.csv("data-sets/VALD - Dynamo WBB.csv")
Forcedecks_B <- read.csv("data-sets/VALD - ForceDecks WBB.csv")
Performance_B <- read.csv("data-sets/VALD - Performance Test WBB.csv")

#womens soccer datasets
Incident_S <- read.csv("data-sets/Incident Report WSOC.csv")
Dynamo_S <- read.csv("data-sets/VALD - Dynamo WSOC.csv")
Forcedecks_S <- read.csv("data-sets/VALD - ForceDecks WSOC.csv")
Performance_S <- read.csv("data-sets/VALD - Performance Test WSOC.csv")

#womens lacrosse datasets
Incident_L <- read.csv("data-sets/Incident Report WLAX.csv")
Dynamo_L <- read.csv("data-sets/VALD - Dynamo WLAX.csv")
Forcedecks_L <- read.csv("data-sets/VALD - ForceDecks WLAX.csv")
Performance_L <- read.csv("data-sets/VALD - Performance Test WLAx (1).csv")

#womens volleyball datasets 
Incident_V <- read.csv("data-sets/Incident Report WVB.csv")
Dynamo_V <- read.csv("data-sets/VALD - Dynamo WVB.csv")
Forcedecks_L <- read.csv("data-sets/VALD - ForceDecks WVB.csv")
Performance_V <- read.csv("data-sets/VALD - Performance Test WVB.csv")
```

```{r}
#copied as paths other way not working rn
#INCIDENT DATASETS
Incident_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WBB.csv")
Incident_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WSOC.csv")
Incident_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WLAX.csv")
Incident_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WVB.csv")
```

#Data Cleaning
-create lower body injury column
-no risk status watchlist to go off of , just look at actual incidents 

1. Incident Datasets
```{r}
#BASKETBALL Incident dataset
Incident_B <- Incident_B %>%                    #189 -> 128 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_B <- subset(Incident_B,select=c("anon_id","Date", "Position", "Incident.Type", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism", "Month", "Year","Incident.Type","Status", "Season", "Start.of.Status", "End.Date"))  #add "Specific.Mechanism","Injury.Position",
        # 128 obs. + 19 vars.

#rougly 4 months of data

#Lower body injuries 
lower_body <- c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip")
incident_B$LB.Injury <- ifelse(incident_B$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0) #lower body Injury Binary
sum(incident_B$LB.Injury)

#ACL injury (most intense LB injury)
incident_B$ACL <- ifelse(incident_B$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_B$ACL)
#HSI injury (overuse injury)
incident_B$HSI <- ifelse(incident_B$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_B$HSI)
```
In womens basketball we have 80 total lower body injuries, 3 of those injuries are ACL injuries, and none are HSI injuries.

```{r}
#CARTERS TO INCORPORATE
#incident_b <- subset(Incident_B, select = c('anon_id', 'Date.of.Injury...Onset.of.symptoms', 'Body.Part', 'OSICS14.Code', 'Final.Diagnosis', 'Recurrence.of.Injury', 'Total.Time.Injured'))

incident_b <- incident_b %>%
  filter(Body.Part %in% lower_body) %>%
  group_by(anon_id, Date.of.Injury...Onset.of.symptoms) %>%
  mutate(Total.Time.Injured = sum(Total.Time.Injured, na.rm = TRUE),
         Date = as.Date(Date.of.Injury...Onset.of.symptoms, format = '%m/%d/%Y')) %>%
  ungroup() %>%
  group_by(across(everything())) %>%
  summarise(.groups = 'drop') %>%
  dplyr::select(-Date.of.Injury...Onset.of.symptoms)
```

```{r}
#SOCCER INCIDENT dataset
Incident_S <- Incident_S %>%                    #361 -> 288 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_S <- subset(Incident_S,select=c("anon_id","Date", "Position", "Incident.Type", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Injured.While.", "Month", "Year","Incident.Type","Status", "Season", "Start.of.Status", "End.Date"))
        #288 obs. + 19 vars.

#roughly 4 months of data / sept. 9 to jan. 15

#Lower body injuriy Binary
incident_S$LB.Injury <- ifelse(incident_S$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_S$LB.Injury)

#ACL injury (most intense LB injury)
incident_S$ACL <- ifelse(incident_S$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_S$ACL)
#HSI injury (overuse injury)
incident_S$HSI <- ifelse(incident_S$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_S$HSI)
```
In women's soccer there is a total of 205 lower body injuries, with 4 of them being ACL injuries, and 7 being HSI injuries.

```{r}
#VOLLEYBALL incident dataset
Incident_V <- Incident_V %>%                    #220 -> 135 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_V <- subset(Incident_V,select=c("anon_id","Date", "Position", "Incident.Type", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Injured.While.", "Month", "Year","Incident.Type","Status", "Season", "Start.of.Status", "End.Date"))
        #135 +23 vars.

#1/16/24-9/8/24 (9 MONTHS)

#Lower body injuriy Binary
incident_V$LB.Injury <- ifelse(incident_V$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_V$LB.Injury)

#ACL injury (most intense LB injury)
incident_V$ACL <- ifelse(incident_V$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_V$ACL)
#HSI injury (overuse injury)
incident_V$HSI <- ifelse(incident_V$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_V$HSI)
```
In womens' volleyball 60 lower body injuries occured, 10 of them beig ACL injuries, 4 of them being HSI injuries. 


```{r}
#LACROSSE incident dataset 
Incident_L <- Incident_L %>%                    #139 -> 86 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_L <- subset(Incident_L,select=c("anon_id","Date", "Position", "Incident.Type", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Injured.While.", "Month", "Year","Incident.Type","Status", "Season", "Start.of.Status", "End.Date"))

#9/9/24-1/13/25 (4.5 MONTHS)

#Lower body injuiy Binary
incident_L$LB.Injury <- ifelse(incident_L$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_L$LB.Injury)

#ACL injury (most intense LB injury)
incident_L$ACL <- ifelse(incident_L$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_L$ACL)
#HSI injury (overuse injury)
incident_L$HSI <- ifelse(incident_L$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_L$HSI)
```
In womens lacrosse 64 lower body injuries occurred, 8 being ACL related, 8 being HSI injury related.




```{r}
#selects variables of interest
dynamo_b <- subset(Dynamo_B, select = c('anon_id', 'Date', 'Knee.Ex.Right.Side.Max.Force..N.', 'Knee.Ex.Left.Side.Max.Force..N.', 'Knee.Ex.Max.Force..N..Asymmetry', 'Knee.Ex.Right.Side.Avg.Force..N.', 'Knee.Ex.Left.Side.Avg.Force..N.', 'Knee.Ex.Avg.Force.Asymmetry', 'Knee.Ex.Right.Side.Max.Impulse..Ns.', 'Knee.Ex.Left.Side.Max.Impulse..Ns.', 'Knee.Ex.Max.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Avg.Impulse..Ns.', 'Knee.Ex.Left.Side.Avg.Impulse..Ns.', 'Knee.Ex.Avg.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Max.Rate.of.Force.Development.Asymmetry', 'Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry'))

#converts the date to a date object and omits the NAs
dynamo_b <- dynamo_b %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()
```

There are three NA values in Knee.Ex.Max.Force..N..Asymmetry. Not sure how to calculate this value. 

2. VALD- Dynamo datasets


3. VALD- Forcedecks datasets
```{r}
Forcedecks_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WBB.csv")
Forcedecks_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WSOC.csv")
Forcedecks_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WLAX.csv")
Forcedecks_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WVB.csv")
```

```{r}
#forcedecks (VALD) bball
#BASKETBALL Incident dataset
Forcedecks_B <- Forcedecks_B %>%                    #189 -> 128 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_B <- subset(Forcedecks_B,select=c("anon_id","Date", ))
        # 128 obs. + 19 vars.

#rougly 4 months of data
```

```{r}
#forcedecks (VALD) soccer

```

```{r}
#forcedecks (VALD) lacrosse 

```

```{r}
#forcedecks (VALD) volleyball

```


4. VALD- Performance Test datasets
```{r}
Performance_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WBB.csv")
Performance_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WSOC.csv")
Performance_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WLAx (1).csv")
Performance_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WVB.csv")
```




## Exploratory Analysis 

#Q1: What are meaningful thresholds for strength metrics as they relate to lower body injury risk? 
a. For example, one of our strength coaches uses 5 times body weight (kg) in Newtons for hamstring strength as a cutoff for injury risk. Is this supported by our data? 

-looking for meaningful strength metrics to lower body injury risk
-create thresholds fir these metrics by each position in each sport

#Q2:What trends in strength do we see across the women’s team sports? 
a. Soccer, lacrosse, volleyball, basketball. 



