---
title: "sydneys_markdown"
output: html_document
date: "2025-07-14"
---

```{r}
#Authors: [Sydney S]
#Author Date: 07/09/2025
#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 07/16/2025

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(readr)
```

#Loading Datasets
```{r}
#Womens basketball datasets
Incident_B <- read.csv("data-sets/Incident Report WBB.csv")
Dynamo_B <- read.csv("data-sets/VALD - Dynamo WBB.csv")
Forcedecks_B <- read.csv("data-sets/VALD - ForceDecks WBB.csv")
Performance_B <- read.csv("data-sets/VALD - Performance Test WBB.csv")

#womens soccer datasets
Incident_S <- read.csv("data-sets/Incident Report WSOC.csv")
Dynamo_S <- read.csv("data-sets/VALD - Dynamo WSOC.csv")
Forcedecks_S <- read.csv("data-sets/VALD - ForceDecks WSOC.csv")
Performance_S <- read.csv("data-sets/VALD - Performance Test WSOC.csv")

#womens lacrosse datasets
Incident_L <- read.csv("data-sets/Incident Report WLAX.csv")
Dynamo_L <- read.csv("data-sets/VALD - Dynamo WLAX.csv")
Forcedecks_L <- read.csv("data-sets/VALD - ForceDecks WLAX.csv")
Performance_L <- read.csv("data-sets/VALD - Performance Test WLAx (1).csv")

#womens volleyball datasets 
Incident_V <- read.csv("data-sets/Incident Report WVB.csv")
Dynamo_V <- read.csv("data-sets/VALD - Dynamo WVB.csv")
Forcedecks_L <- read.csv("data-sets/VALD - ForceDecks WVB.csv")
Performance_V <- read.csv("data-sets/VALD - Performance Test WVB.csv")
```

```{r}
#copied as paths above were not working rn
#INCIDENT DATASETS
Incident_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WBB.csv")
Incident_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WSOC.csv")
Incident_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WLAX.csv")
Incident_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/Incident Report WVB.csv")
```

#Data Cleaning
-no risk status watchlist to compare to, just look at actual incidents 

1. Incident Datasets
```{r}
#BASKETBALL Incident dataset
Incident_B <- Incident_B %>%                    #189 -> 128 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") 
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_B <- subset(Incident_B,select=c("anon_id","Date", "Position", "Date.of.Injury...Onset.of.symptoms", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis","Final.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Specific.Mechanism","Injured.Position", "Month", "Year","Status","Season.","Start.of.Status", "End.Date"))  #add,"Injured.Position",
        # 128 obs. + 26 vars.

#Lower body injuries 
lower_body <- c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip")
incident_B$LB.Injury <- ifelse(incident_B$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0) #lower body Injury Binary
sum(incident_B$LB.Injury)

#ACL injury (most intense LB injury)
incident_B$ACL <- ifelse(incident_B$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_B$ACL)
#HSI injury (overuse injury)
incident_B$HSI <- ifelse(incident_B$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_B$HSI)

#12/5/23-1/2/25 DATE RANGE (2 years)

#converting all dates
incident_b <- incident_B %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
       Start.of.Status=as.Date(Start.of.Status,format='%m/%d/%Y'),
       End.Date= as.Date(End.Date, format='%m/%d/%Y'),
       Date.of.Injury...Onset.of.symptoms= as.Date(Date.of.Injury...Onset.of.symptoms, format='%m/%d/%Y'))

#create a length of status variable (Out, limited,as tolerated, full go)
incident_b$Length.of.Status <- difftime(incident_b$End.Date, incident_b$Start.of.Status,units=c("days"))
```
In womens basketball we have 80 total lower body injuries, 3 of those injuries are ACL injuries, and none are HSI injuries.
```{r}
#bball incidents plots
#position by count if lower body injury 

```

```{r}
#CARTERS TO INCORPORATE LATER
incident_b <- incident_b %>%
  filter(Body.Part %in% lower_body) %>%
  group_by(anon_id, Date.of.Injury...Onset.of.symptoms) %>%
  mutate(Total.Time.Injured = sum(Total.Time.Injured, na.rm = TRUE),
         Date = as.Date(Date.of.Injury...Onset.of.symptoms, format = '%m/%d/%Y')) %>%
  ungroup() %>%
  group_by(across(everything())) %>%
  summarise(.groups = 'drop') %>%
  dplyr::select(-Date.of.Injury...Onset.of.symptoms)
```

```{r}
#SOCCER INCIDENT dataset
Incident_S <- Incident_S %>%                    #361 -> 288 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_S <- subset(Incident_S,select=c("anon_id","Date", "Position", "Date.of.Injury...Onset.of.symptoms", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Final.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Specific.Mechanism","Injured.While.", "Month", "Year","Incident.Type","Status", "Season.","Start.of.Status", "End.Date"))
        #288 obs. + 27 vars.

#Lower body injuriy Binary
incident_S$LB.Injury <- ifelse(incident_S$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_S$LB.Injury)

#ACL injury (most intense LB injury)
incident_S$ACL <- ifelse(incident_S$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_S$ACL)
#HSI injury (overuse injury)
incident_S$HSI <- ifelse(incident_S$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_S$HSI)

#5/8/23- 1/15/25 (year and a half ish)

#converting all dates
incident_s <- incident_S %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
       Start.of.Status=as.Date(Start.of.Status,format='%m/%d/%Y'),
       End.Date= as.Date(End.Date, format='%m/%d/%Y'),
       Date.of.Injury...Onset.of.symptoms= as.Date(Date.of.Injury...Onset.of.symptoms, format='%m/%d/%Y'))

#create a length of status variable (Out, limited,as tolerated, full go)
incident_s$Length.of.Status <- difftime(incident_s$End.Date, incident_s$Start.of.Status,units=c("days"))
```
In women's soccer there is a total of 205 lower body injuries, with 4 of them being ACL injuries, and 7 being HSI injuries.
```{r}
#soccer incident plot
#position by count of lower body injuries 
```

```{r}
#VOLLEYBALL incident dataset
Incident_V <- Incident_V %>%                    #220 -> 135 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_V <- subset(Incident_V,select=c("anon_id","Date", "Position", "Date.of.Injury...Onset.of.symptoms", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Final.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Specific.Mechanism","Injured.While.", "Month", "Year","Status", "Season.","Start.of.Status", "End.Date"))
        #135 +23 vars.

#Lower body injuriy Binary
incident_V$LB.Injury <- ifelse(incident_V$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_V$LB.Injury)

#ACL injury (most intense LB injury)
incident_V$ACL <- ifelse(incident_V$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_V$ACL)
#HSI injury (overuse injury)
incident_V$HSI <- ifelse(incident_V$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_V$HSI)

#1/16/24-9/8/24 (9 MONTHS)

#converting all dates
incident_v <- incident_V %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
       Start.of.Status=as.Date(Start.of.Status,format='%m/%d/%Y'),
       End.Date= as.Date(End.Date, format='%m/%d/%Y'),
       Date.of.Injury...Onset.of.symptoms= as.Date(Date.of.Injury...Onset.of.symptoms, format='%m/%d/%Y'))

#create a length of status variable (Out, limited,as tolerated, full go)
incident_v$Length.of.Status <- difftime(incident_v$End.Date, incident_v$Start.of.Status,units=c("days"))
```
In womens' volleyball 60 lower body injuries occured, 10 of them beig ACL injuries, 4 of them being HSI injuries. 
```{r}
#vball incident plot
#positions by count of lower body injuries 
```


```{r}
#LACROSSE incident dataset 
Incident_L <- Incident_L %>%                    #139 -> 86 obs.
filter(Incident.Type=="Injury") %>%
filter(Result.of.Sport.Participation=="Yes") #%>%
#filter(General.mechanism=="Non-contact"|"Chronic/Overuse")

incident_L <- subset(Incident_L,select=c("anon_id","Date", "Position", "Date.of.Injury...Onset.of.symptoms", "Side", "Body.Part.", "Tissue.Type","OSICS14.Code","OSICS10.Diagnosis", "Final.Diagnosis", "Result.of.Sport.Participation", "Recurrence.of.Injury", "Injury.Prognosis", "General.mechanism","Specific.Mechanism", "Injured.While.", "Month", "Year","Incident.Type","Status", "Season.","Start.of.Status", "End.Date"))

#Lower body injuiy Binary
incident_L$LB.Injury <- ifelse(incident_L$Body.Part. %in% c("Thigh", "Ankle", "Lower Leg", "Knee", "Foot", "Groin/hip"),1,0)
sum(incident_L$LB.Injury)

#ACL injury (most intense LB injury)
incident_L$ACL <- ifelse(incident_L$OSICS14.Code %in% c("KL1","KLA"),1,0)
sum(incident_L$ACL)
#HSI injury (overuse injury)
incident_L$HSI <- ifelse(incident_L$OSICS14.Code %in% c("TM1","TMB","TMS"),1,0)
sum(incident_L$HSI)

#4/12/23-4/18/25 (2 years)

#converting all dates
incident_l <- incident_L %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
       Start.of.Status=as.Date(Start.of.Status,format='%m/%d/%Y'),
       End.Date= as.Date(End.Date, format='%m/%d/%Y'),
       Date.of.Injury...Onset.of.symptoms= as.Date(Date.of.Injury...Onset.of.symptoms, format='%m/%d/%Y'))

#create a length of status variable (Out, limited,as tolerated, full go)
incident_l$Length.of.Status <- difftime(incident_l$End.Date, incident_l$Start.of.Status,units=c("days"))
```
In womens lacrosse 64 lower body injuries occurred, 8 being ACL related, 8 being HSI injury related.


2. VALD- Dynamo datasets
-we need to know the calculations for the asymmetry measures because they seem inconsistent.
```{r}
#load data
Dynamo_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Dynamo WBB.csv")
Dynamo_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Dynamo WSOC.csv")
Dynamo_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Dynamo WVB.csv")
Dynamo_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Dynamo WLAX.csv")
```

**1. Basketball**
```{r}
#selects variables of interest
dynamo_B <- subset(Dynamo_B, select = c('anon_id', 'Date', 'Knee.Ex.Right.Side.Max.Force..N.', 'Knee.Ex.Left.Side.Max.Force..N.', 'Knee.Ex.Max.Force..N..Asymmetry', 'Knee.Ex.Right.Side.Avg.Force..N.', 'Knee.Ex.Left.Side.Avg.Force..N.', 'Knee.Ex.Avg.Force.Asymmetry', 'Knee.Ex.Right.Side.Max.Impulse..Ns.', 'Knee.Ex.Left.Side.Max.Impulse..Ns.', 'Knee.Ex.Max.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Avg.Impulse..Ns.', 'Knee.Ex.Left.Side.Avg.Impulse..Ns.', 'Knee.Ex.Avg.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Max.Rate.of.Force.Development.Asymmetry', 'Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry'))

#converts the date to a date object and omits the NAs
dynamo_B <- dynamo_B %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
         Knee.Ex.Max.Force..N..Asymmetry = case_when(
           Knee.Ex.Left.Side.Max.Force..N. > Knee.Ex.Right.Side.Max.Force..N. ~ round(100 * (1 - (Knee.Ex.Right.Side.Max.Force..N. / Knee.Ex.Left.Side.Max.Force..N.)), 1),
           TRUE ~ Knee.Ex.Max.Force..N..Asymmetry
         )) %>%
  na.omit()
```
19 total observations, 3 with NA values in Knee.Ex.Max.Force..N..Asymmetry.

**2. Soccer**
```{r, warning = FALSE}
#selects variables of interest
dynamo_S <- subset(Dynamo_S, select = c('anon_id', 'Date', 'Knee.Ex.Right.Side.Max.Force..N.', 'Knee.Ex.Left.Side.Max.Force..N.', 'Knee.Ex.Max.Force..N..Asymmetry', 'Knee.Ex.Right.Side.Avg.Force..N.', 'Knee.Ex.Left.Side.Avg.Force..N.', 'Knee.Ex.Avg.Force.Asymmetry', 'Knee.Ex.Right.Side.Max.Impulse..Ns.', 'Knee.Ex.Left.Side.Max.Impulse..Ns.', 'Knee.Ex.Max.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Avg.Impulse..Ns.', 'Knee.Ex.Left.Side.Avg.Impulse..Ns.', 'Knee.Ex.Avg.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Max.Rate.of.Force.Development.Asymmetry', 'Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry'))

#converts the date to a date object and puts values from different observations together into one observation
dynamo_S <- dynamo_S %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date) %>%
  summarise(Knee.Ex.Right.Side.Max.Force..N. = sum(Knee.Ex.Right.Side.Max.Force..N., na.rm = TRUE), 
            Knee.Ex.Left.Side.Max.Force..N. = sum(Knee.Ex.Left.Side.Max.Force..N., na.rm = TRUE), 
            Knee.Ex.Max.Force..N..Asymmetry = sum(Knee.Ex.Max.Force..N..Asymmetry), 
            Knee.Ex.Right.Side.Avg.Force..N. = sum(Knee.Ex.Right.Side.Avg.Force..N., na.rm = TRUE), 
            Knee.Ex.Left.Side.Avg.Force..N. = sum(Knee.Ex.Left.Side.Avg.Force..N., na.rm = TRUE), 
            Knee.Ex.Avg.Force.Asymmetry = sum(Knee.Ex.Avg.Force.Asymmetry), 
            Knee.Ex.Right.Side.Max.Impulse..Ns. = sum(Knee.Ex.Right.Side.Max.Impulse..Ns., na.rm = TRUE), 
            Knee.Ex.Left.Side.Max.Impulse..Ns. = sum(Knee.Ex.Left.Side.Max.Impulse..Ns., na.rm = TRUE), 
            Knee.Ex.Max.Impulse.Asymmetry = sum(Knee.Ex.Max.Impulse.Asymmetry), 
            Knee.Ex.Right.Side.Avg.Impulse..Ns. = sum(Knee.Ex.Right.Side.Avg.Impulse..Ns., na.rm = TRUE), 
            Knee.Ex.Left.Side.Avg.Impulse..Ns. = sum(Knee.Ex.Left.Side.Avg.Impulse..Ns., na.rm = TRUE), 
            Knee.Ex.Avg.Impulse.Asymmetry = sum(Knee.Ex.Avg.Impulse.Asymmetry),
            Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns. = sum(Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns., na.rm = TRUE),
            Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns. = sum(Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns., na.rm = TRUE),
            Knee.Ex.Max.Rate.of.Force.Development.Asymmetry = sum(Knee.Ex.Max.Rate.of.Force.Development.Asymmetry),
            Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns. = sum(Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns., na.rm = TRUE),
            Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns. = sum(Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns., na.rm = TRUE),
            Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry = sum(Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry), .groups = "drop") %>%
  mutate(Knee.Ex.Max.Force..N..Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Max.Force..N. / Knee.Ex.Right.Side.Max.Force..N.)), 1),
         Knee.Ex.Avg.Force.Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Avg.Force..N. / Knee.Ex.Right.Side.Avg.Force..N.)), 1),
         Knee.Ex.Max.Impulse.Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Max.Impulse..Ns. / Knee.Ex.Right.Side.Max.Impulse..Ns.)), 1),
         Knee.Ex.Avg.Impulse.Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Avg.Impulse..Ns. / Knee.Ex.Right.Side.Avg.Impulse..Ns.)), 1),
         Knee.Ex.Max.Rate.of.Force.Development.Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns. / Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.)), 1),
         Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry = round(100 * (1 - (Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns. / Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.)), 1))
```
Only 1 observation, none of the asymmetry statistics are summarized. 

**3. Lacrosse**
```{r}
#selects variables of interest
dynamo_L <- subset(Dynamo_L, select = c('anon_id', 'Date', 'Knee.Ex.Right.Side.Max.Force..N.', 'Knee.Ex.Left.Side.Max.Force..N.', 'Knee.Ex.Max.Force..N..Asymmetry', 'Knee.Ex.Right.Side.Avg.Force..N.', 'Knee.Ex.Left.Side.Avg.Force..N.', 'Knee.Ex.Avg.Force.Asymmetry', 'Knee.Ex.Right.Side.Max.Impulse..Ns.', 'Knee.Ex.Left.Side.Max.Impulse..Ns.', 'Knee.Ex.Max.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Avg.Impulse..Ns.', 'Knee.Ex.Left.Side.Avg.Impulse..Ns.', 'Knee.Ex.Avg.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Max.Rate.of.Force.Development.Asymmetry', 'Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry'))

#converts the date to a date object and omits the NAs
dynamo_L <- dynamo_L %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
         Knee.Ex.Max.Force..N..Asymmetry = case_when(
           Knee.Ex.Left.Side.Max.Force..N. > Knee.Ex.Right.Side.Max.Force..N. ~ round(100 * (1 - (Knee.Ex.Right.Side.Max.Force..N. / Knee.Ex.Left.Side.Max.Force..N.)), 1),
           TRUE ~ Knee.Ex.Max.Force..N..Asymmetry
         )) %>%
  na.omit()
```

**4. Volleyball**
```{r}
#selects variables of interest
dynamo_V <- subset(Dynamo_V, select = c('anon_id', 'Date', 'Knee.Ex.Right.Side.Max.Force..N.', 'Knee.Ex.Left.Side.Max.Force..N.', 'Knee.Ex.Max.Force..N..Asymmetry', 'Knee.Ex.Right.Side.Avg.Force..N.', 'Knee.Ex.Left.Side.Avg.Force..N.', 'Knee.Ex.Avg.Force.Asymmetry', 'Knee.Ex.Right.Side.Max.Impulse..Ns.', 'Knee.Ex.Left.Side.Max.Impulse..Ns.', 'Knee.Ex.Max.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Avg.Impulse..Ns.', 'Knee.Ex.Left.Side.Avg.Impulse..Ns.', 'Knee.Ex.Avg.Impulse.Asymmetry', 'Knee.Ex.Right.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Max.Rate.of.Force.Development..Ns.', 'Knee.Ex.Max.Rate.of.Force.Development.Asymmetry', 'Knee.Ex.Right.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Left.Side.Avg.Rate.of.Force.Development..Ns.', 'Knee.Ex.Avg.Rate.of.Force.Development.Asymmetry'))

#converts the date to a date object and omits the NAs
dynamo_V <- dynamo_V %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y'),
         Knee.Ex.Max.Force..N..Asymmetry = case_when(
           Knee.Ex.Left.Side.Max.Force..N. > Knee.Ex.Right.Side.Max.Force..N. ~ round(100 * (1 - (Knee.Ex.Right.Side.Max.Force..N. / Knee.Ex.Left.Side.Max.Force..N.)), 1),
           TRUE ~ Knee.Ex.Max.Force..N..Asymmetry
         )) %>%
  na.omit()
```
Only has 4 observations from one player but not missing any data.




3. VALD- Forcedecks datasets
```{r}
#import datasets
Forcedecks_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WBB.csv")
Forcedecks_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WSOC.csv")
Forcedecks_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WLAX.csv")
Forcedecks_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - ForceDecks WVB.csv")
```

#Forcedecks (VALD) basketball data
```{r}
#TEST TYPE=CMJ
CMJ_B <- subset(Forcedecks_B, select = c("anon_id","Date","Position","Test.Type","Weight..kg.","Countermovement.Depth..cm.","Peak.Power...BM..W.kg.","Landing.RFD..N.s.", "Eccentric.Concentric.Mean.Force.Ratio", "Peak.Landing.Force..Asym...N.", "Peak.Power..W.","Concentric.Maximum.RFD..N.s.","Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.","Eccentric.Mean.Braking.Force..N.", "Max.Jump.Height..cm.", "Max.Concentric.Peak.Force..N.", "Relative.Concentric.Power.Bilateral","Relative.Eccentric.Power.Bilateral", "Max.Peak.Power...BM", "Max.Eccentric.Peak.Force..N.","Max.RSI.Modified.VALD","Max.Concentric.RPD..N.s.","Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "Max.Peak.Landing.Force..N.", "MAX.Concentric.Peak.Power..W.", "Max.CMJ.Jump.Height..cm.","Max.CMJ.RSI.Modified..m.s.", "Max.CMJ.Concentric.RPD..N.s.", "Max.CMJ.Peak.Power..W.", "Max.CMJ.Peak.Landing.Force..N.","CMJ.Concentric.Relative.Peak.Power...Difference.from.First","CMJ.Eccentric.Relative.Peak.Power...Difference.from.First","Eccentric.Asymmetry","Concentric.Asymmetry"
))            #2751 obs & 36 vars.

#possible other final dataset 
cmj_b <- CMJ_B %>%
  na.omit()   #440 obs

#convert empty vectors to NA
CMJ_B[CMJ_B==""] <- NA    #colSums(is.na(forcedecks_B)) look at NAs in columns

#to use in summarise
#function for mode
get_mode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

#set all needed variables to numeric
CMJ_B$Countermovement.Depth..cm. <- as.numeric(CMJ_B$Countermovement.Depth..cm.)
CMJ_B$Peak.Power...BM..W.kg. <- as.numeric(CMJ_B$Peak.Power...BM..W.kg.)

#combining values from different observations into one
CMJ_b <- CMJ_B %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date,Test.Type) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Countermovement.Depth..cm.= mean(Countermovement.Depth..cm., na.rm=TRUE),
            Peak.Power...BM..W.kg.= mean(Peak.Power...BM..W.kg., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s.,na.rm=TRUE),
            Eccentric.Concentric.Mean.Force.Ratio= mean(Eccentric.Concentric.Mean.Force.Ratio,na.rm=TRUE),
            Peak.Landing.Force..Asym...N.= mean(Peak.Landing.Force..Asym...N.,na.rm=TRUE),
            Peak.Power..W.= mean(Peak.Power..W.,na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Relative.Concentric.Power.Bilateral= mean(Relative.Concentric.Power.Bilateral,na.rm=TRUE),
            Relative.Eccentric.Power.Bilateral= mean(Relative.Eccentric.Power.Bilateral,na.rm=TRUE),
            Max.Peak.Power...BM= mean(Max.Peak.Power...BM,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD,na.rm=TRUE),
            Max.Concentric.RPD..N.s.= mean(Max.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Peak.Landing.Force..N.= mean(Max.Peak.Landing.Force..N.,na.rm=TRUE),
            MAX.Concentric.Peak.Power..W.= mean(MAX.Concentric.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Jump.Height..cm.= mean(Max.CMJ.Jump.Height..cm.,na.rm=TRUE),
            Max.CMJ.RSI.Modified..m.s.= mean(Max.CMJ.RSI.Modified..m.s.,na.rm=TRUE),
            Max.CMJ.Concentric.RPD..N.s.= mean(Max.CMJ.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.CMJ.Peak.Power..W.= mean(Max.CMJ.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Peak.Landing.Force..N.= mean(Max.CMJ.Peak.Landing.Force..N.,na.rm=TRUE),
            CMJ.Concentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Concentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            CMJ.Eccentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Eccentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            Position= get_mode(Position),
            Eccentric.Asymmetry= get_mode(Eccentric.Asymmetry),
            Concentric.Asymmetry= get_mode(Concentric.Asymmetry),
            .groups = 'drop'
              )                   #1393 obs & 36 vars
#fix NaN values

CMJ_bb <- CMJ_b %>%
  na.omit()   #196 obs & 36 vars
```

```{r}
#2 TEST TYPE=HJ
HJ_B <- Forcedecks_B %>%
  filter(Test.Type=="HJ")  %>% #424 obs
  select("anon_id", "Date", "Corrected.Standing.Weight..Kg.", "Mean.RSI..Flight.Contact.Time.", "Active.Stiffness..N.m.", "Best.Average.Force..N.", "Best.Average.Force..Asym...N.", "Best.Time.to.Peak.Force..ms.", "Best.Contact.Time..ms.", "Best.Flight.Time..ms.", "Best.Impulse..N.s.", "Best.Peak.Force..N.", "Best.Peak.Force..Asym...N.", "Mean.Average.Force..N.", "Mean.Average.Force..Asym...N.", "Mean.Landing.RFD..N.s.", "Mean.Landing.RFD..Asym...N.s.", "Stiffness.Fatigue....", "Stiffness.Fatigue..Asym.....", "Max.Jump.Height..cm.") %>%
  na.omit()
                 #409 obs & 20 vars
```

```{r}
#3: TEST TYPE= IMTP
IMTP_B <- Forcedecks_B %>%
  filter(Test.Type=="IMTP") %>% #162 
  select("anon_id", "Date", "Baseline.Force..N.","Baseline.Force..Asym...N.", "Peak.Vertical.Force..N.", "Peak.Vertical.Force...BM..N.kg.", "Peak.Vertical.Force..Asym...N.", "RFD...50ms..N.s.", "RFD...50ms..Asym...N.s.", "RFD...100ms..N.s.", "RFD...100ms..Asym...N.s.", "RFD...150ms..N.s.", "RFD...150ms..Asym...N.s.", "RFD...200ms..N.s.", "RFD...200ms..Asym...N.s.", "Net.Peak.Vertical.Force..N.s.", "Net.Peak.Vertical.Force..Asym...N.s.", "Force.at.200ms...BM..N.kg.", "Eccentric.Asymmetry", "Concentric.Asymmetry","MEAN.Eccentric.Mean.Force.Left", "MEAN.Eccentric.Mean.Force.Right", "MEAN.Concentric.Mean.Force.Left", "MEAN.Concentric.Mean.Force.Right", "MEAN.Concentric.Mean.Force.Asym", "MEAN.Peak.Landing.Force.Asym",  "MEAN.Relative.Peak.Power", "MEAN.CMJ.Concentric.Force", "MEAN.Peak.Landing.Force.Left", "MEAN.Peak.Landing.Force.Right") %>%
  na.omit()   #50 obs & 30 vars
```

#Forcedecks (VALD) Soccer Data
```{r}
#1: TEST TYPE==CMJ
#selects variables of interest
CMJ_S <- subset(Forcedecks_S, select = c("anon_id","Date","Position","Test.Type","Weight..kg.","Countermovement.Depth..cm.","Peak.Power...BM..W.kg.","Landing.RFD..N.s.", "Eccentric.Concentric.Mean.Force.Ratio", "Peak.Landing.Force..Asym...N.", "Peak.Power..W.","Concentric.Maximum.RFD..N.s.","Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.","Eccentric.Mean.Braking.Force..N.", "Max.Jump.Height..cm.", "Max.Concentric.Peak.Force..N.", "Relative.Concentric.Power.Bilateral",
"Relative.Eccentric.Power.Bilateral","Max.Peak.Power...BM", "Max.Eccentric.Peak.Force..N.",
"Max.RSI.Modified.VALD","Max.Concentric.RPD..N.s.", "Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "Max.Peak.Landing.Force..N.", "MAX.Concentric.Peak.Power..W.","Max.CMJ.Jump.Height..cm.",
"Max.CMJ.RSI.Modified..m.s.","Max.CMJ.Concentric.RPD..N.s.","Max.CMJ.Peak.Power..W.","Max.CMJ.Peak.Landing.Force..N.","CMJ.Concentric.Relative.Peak.Power...Difference.from.First","CMJ.Eccentric.Relative.Peak.Power...Difference.from.First","Eccentric.Asymmetry","Concentric.Asymmetry"
))        #3443 obs & 36 vars 

#OTHER POSSIBLE FINAL DATASET
cmj_s <- CMJ_S %>%
na.omit()  #562 obs & 36 vars
cmj_s[cmj_s==""] <- NA

CMJ_S[CMJ_S==""] <- NA

#combining values from different observations into one
CMJ_s <- CMJ_S %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date, Test.Type) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Countermovement.Depth..cm.= mean(Countermovement.Depth..cm., na.rm=TRUE),
            Peak.Power...BM..W.kg.= mean(Peak.Power...BM..W.kg., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s.,na.rm=TRUE),
            Eccentric.Concentric.Mean.Force.Ratio= mean(Eccentric.Concentric.Mean.Force.Ratio,na.rm=TRUE),
            Peak.Landing.Force..Asym...N.= mean(Peak.Landing.Force..Asym...N.,na.rm=TRUE),
            Peak.Power..W.= mean(Peak.Power..W.,na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Relative.Concentric.Power.Bilateral= mean(Relative.Concentric.Power.Bilateral,na.rm=TRUE),
            Relative.Eccentric.Power.Bilateral= mean(Relative.Eccentric.Power.Bilateral,na.rm=TRUE),
            Max.Peak.Power...BM= mean(Max.Peak.Power...BM,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD,na.rm=TRUE),
            Max.Concentric.RPD..N.s.= mean(Max.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Peak.Landing.Force..N.= mean(Max.Peak.Landing.Force..N.,na.rm=TRUE),
            MAX.Concentric.Peak.Power..W.= mean(MAX.Concentric.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Jump.Height..cm.= mean(Max.CMJ.Jump.Height..cm.,na.rm=TRUE),
            Max.CMJ.RSI.Modified..m.s.= mean(Max.CMJ.RSI.Modified..m.s.,na.rm=TRUE),
            Max.CMJ.Concentric.RPD..N.s.= mean(Max.CMJ.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.CMJ.Peak.Power..W.= mean(Max.CMJ.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Peak.Landing.Force..N.= mean(Max.CMJ.Peak.Landing.Force..N.,na.rm=TRUE),
            CMJ.Concentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Concentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            CMJ.Eccentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Eccentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            Position= get_mode(Position),
            Eccentric.Asymmetry= get_mode(Eccentric.Asymmetry),
            Concentric.Asymmetry= get_mode(Concentric.Asymmetry),
            .groups = 'drop'
              )               # 1077 obs & 32 vars -> 1571

CMJ_soc <- CMJ_s %>%
  na.omit()   #621 obs & 32 vars -> 600 -> 257 obs & 36 vars 
```

```{r}
#2 TEST TYPE= HJ
HJ_S <- Forcedecks_S %>%
  filter(Test.Type=="HJ")  %>%  #575 obs
  select("anon_id", "Date", "Corrected.Standing.Weight..Kg.", "Mean.RSI..Flight.Contact.Time.", "Active.Stiffness..N.m.", "Best.Average.Force..N.", "Best.Average.Force..Asym...N.", "Best.Time.to.Peak.Force..ms.", "Best.Contact.Time..ms.", "Best.Flight.Time..ms.", "Best.Impulse..N.s.", "Best.Peak.Force..N.", "Best.Peak.Force..Asym...N.", "Mean.Average.Force..N.", "Mean.Average.Force..Asym...N.", "Mean.Landing.RFD..N.s.", "Mean.Landing.RFD..Asym...N.s.", "Stiffness.Fatigue....", "Stiffness.Fatigue..Asym.....", "Max.Jump.Height..cm.") %>%
  na.omit()
                 #556 obs & 20 vars
```

```{r}
#3 TEST TYPE- SLJ
SLJ_S <- Forcedecks_S %>%
  filter(Test.Type=="SLJ")  # 455 obs

#important variables 
SLJ_S <- subset(SLJ_S, select = c("anon_id", "Date", "Position", "Weight..kg.", "Jump.Height..Flight.Time...cm.",  "Jump.Height..Imp.Mom...cm.", "Max.Jump.Height..cm.", "Max.Single.Leg.Jump.Height.Left", "Max.Single.Leg.Jump.Height.Right", "First.SLJ.Jump.Height", "Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "RSI.Modified..m.s.", "Max.RSI.Modified..m.s.", "Max.RSI.Modified.VALD", "Eccentric.Mean.Braking.Force..N.", "Landing.RFD..N.s.","Concentric.Maximum.RFD..N.s.", "Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.", "Concentric.Impulse..Ns.", "Concentric.Impulse..Left...Ns.", "Concentric.Impulse..Right...Ns.", "Max.Concentric.Peak.Force..N.", "Max.Eccentric.Peak.Force..N."
))                        #455 obs & 25 vars

SLJ_S[SLJ_S==""] <- NA

#combining values from different observations into one
SLJ_s <- SLJ_S %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Jump.Height..Flight.Time...cm.= mean(Jump.Height..Flight.Time...cm., na.rm=TRUE),
            Jump.Height..Imp.Mom...cm.= mean(Jump.Height..Imp.Mom...cm.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm. ,na.rm=TRUE),
            Max.Single.Leg.Jump.Height.Left= mean(Max.Single.Leg.Jump.Height.Left ,na.rm=TRUE),
            Max.Single.Leg.Jump.Height.Right= mean(Max.Single.Leg.Jump.Height.Right ,na.rm=TRUE),
            First.SLJ.Jump.Height= mean(First.SLJ.Jump.Height, na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s., na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s., na.rm=TRUE),
            RSI.Modified..m.s.= mean(RSI.Modified..m.s., na.rm=TRUE),
            Max.RSI.Modified..m.s.= mean(Max.RSI.Modified..m.s., na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD, na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s., na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s., na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Concentric.Impulse..Ns.= mean(Concentric.Impulse..Ns.,na.rm=TRUE),
            Concentric.Impulse..Left...Ns.= mean(Concentric.Impulse..Left...Ns.,na.rm=TRUE),
            Concentric.Impulse..Right...Ns.= mean(Concentric.Impulse..Right...Ns.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Position= get_mode(Position),
            .groups = 'drop'
              )   #83 obs 25 vars
```

#Forcedecks (VALD) Lacrosse Data
```{r}
#1 : TEST TYPE= CMJ
#selects variables of interest
CMJ_L <- subset(Forcedecks_L, select = c("anon_id","Date","Position","Test.Type","Weight..kg.","Countermovement.Depth..cm.","Peak.Power...BM..W.kg.","Landing.RFD..N.s.", "Eccentric.Concentric.Mean.Force.Ratio", "Peak.Landing.Force..Asym...N.", "Peak.Power..W.","Concentric.Maximum.RFD..N.s.","Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.","Eccentric.Mean.Braking.Force..N.", "Max.Jump.Height..cm.", "Max.Concentric.Peak.Force..N.", "Relative.Concentric.Power.Bilateral",
"Relative.Eccentric.Power.Bilateral","Max.Peak.Power...BM", "Max.Eccentric.Peak.Force..N.",
"Max.RSI.Modified.VALD","Max.Concentric.RPD..N.s.", "Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "Max.Peak.Landing.Force..N.", "MAX.Concentric.Peak.Power..W.","Max.CMJ.Jump.Height..cm.",
"Max.CMJ.RSI.Modified..m.s.","Max.CMJ.Concentric.RPD..N.s.","Max.CMJ.Peak.Power..W.","Max.CMJ.Peak.Landing.Force..N.","CMJ.Concentric.Relative.Peak.Power...Difference.from.First","CMJ.Eccentric.Relative.Peak.Power...Difference.from.First","Eccentric.Asymmetry","Concentric.Asymmetry"
))        #1939 obs & 36 vars 

#possible other final dataset
cmj_l <- CMJ_L %>%
  na.omit()      #377 obs 

CMJ_L[CMJ_L==""] <- NA

#combining values from different observations into one
CMJ_l <- CMJ_L %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date, Test.Type) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Countermovement.Depth..cm.= mean(Countermovement.Depth..cm., na.rm=TRUE),
            Peak.Power...BM..W.kg.= mean(Peak.Power...BM..W.kg., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s.,na.rm=TRUE),
            Eccentric.Concentric.Mean.Force.Ratio= mean(Eccentric.Concentric.Mean.Force.Ratio,na.rm=TRUE),
            Peak.Landing.Force..Asym...N.= mean(Peak.Landing.Force..Asym...N.,na.rm=TRUE),
            Peak.Power..W.= mean(Peak.Power..W.,na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Relative.Concentric.Power.Bilateral= mean(Relative.Concentric.Power.Bilateral,na.rm=TRUE),
            Relative.Eccentric.Power.Bilateral= mean(Relative.Eccentric.Power.Bilateral,na.rm=TRUE),
            Max.Peak.Power...BM= mean(Max.Peak.Power...BM,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD,na.rm=TRUE),
            Max.Concentric.RPD..N.s.= mean(Max.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Peak.Landing.Force..N.= mean(Max.Peak.Landing.Force..N.,na.rm=TRUE),
            MAX.Concentric.Peak.Power..W.= mean(MAX.Concentric.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Jump.Height..cm.= mean(Max.CMJ.Jump.Height..cm.,na.rm=TRUE),
            Max.CMJ.RSI.Modified..m.s.= mean(Max.CMJ.RSI.Modified..m.s.,na.rm=TRUE),
            Max.CMJ.Concentric.RPD..N.s.= mean(Max.CMJ.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.CMJ.Peak.Power..W.= mean(Max.CMJ.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Peak.Landing.Force..N.= mean(Max.CMJ.Peak.Landing.Force..N.,na.rm=TRUE),
            CMJ.Concentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Concentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            CMJ.Eccentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Eccentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            Position= get_mode(Position),
            Eccentric.Asymmetry= get_mode(Eccentric.Asymmetry),
            Concentric.Asymmetry= get_mode(Concentric.Asymmetry),
            .groups = 'drop'
              )               # 1238 obs & 36 vars

CMJ_lax <- CMJ_l %>%
  na.omit()   #277 obs & 36 vars
```

```{r}
#2: TEST TYPE= HJ
HJ_L <- Forcedecks_L %>%
  filter(Test.Type=="HJ")  %>%  #582 obs
  select("anon_id", "Date", "Corrected.Standing.Weight..Kg.", "Mean.RSI..Flight.Contact.Time.", "Active.Stiffness..N.m.", "Best.Average.Force..N.", "Best.Average.Force..Asym...N.", "Best.Time.to.Peak.Force..ms.", "Best.Contact.Time..ms.", "Best.Flight.Time..ms.", "Best.Impulse..N.s.", "Best.Peak.Force..N.", "Best.Peak.Force..Asym...N.", "Mean.Average.Force..N.", "Mean.Average.Force..Asym...N.", "Mean.Landing.RFD..N.s.", "Mean.Landing.RFD..Asym...N.s.", "Stiffness.Fatigue....", "Stiffness.Fatigue..Asym.....", "Max.Jump.Height..cm.") %>%
  na.omit()
                 #573 obs & 20 vars
```

```{r}
#3: TEST TYPE: SLJ
SLJ_L <- Forcedecks_L %>%
  filter(Test.Type=="SLJ")  #190 obs

#important variables 
SLJ_L <- subset(SLJ_L, select = c("anon_id", "Date", "Position", "Weight..kg.", "Jump.Height..Flight.Time...cm.",  "Jump.Height..Imp.Mom...cm.", "Max.Jump.Height..cm.", "Max.Single.Leg.Jump.Height.Left", "Max.Single.Leg.Jump.Height.Right", "First.SLJ.Jump.Height", "Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "RSI.Modified..m.s.", "Max.RSI.Modified..m.s.", "Max.RSI.Modified.VALD", "Eccentric.Mean.Braking.Force..N.", "Landing.RFD..N.s.","Concentric.Maximum.RFD..N.s.", "Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.", "Concentric.Impulse..Ns.", "Concentric.Impulse..Left...Ns.", "Concentric.Impulse..Right...Ns.", "Max.Concentric.Peak.Force..N.", "Max.Eccentric.Peak.Force..N."
))                        #190 obs & 25 vars

SLJ_L[SLJ_L==""] <- NA

#combining values from different observations into one
SLJ_l <- SLJ_L %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  group_by(anon_id, Date) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Jump.Height..Flight.Time...cm.= mean(Jump.Height..Flight.Time...cm., na.rm=TRUE),
            Jump.Height..Imp.Mom...cm.= mean(Jump.Height..Imp.Mom...cm.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm. ,na.rm=TRUE),
            Max.Single.Leg.Jump.Height.Left= mean(Max.Single.Leg.Jump.Height.Left ,na.rm=TRUE),
            Max.Single.Leg.Jump.Height.Right= mean(Max.Single.Leg.Jump.Height.Right ,na.rm=TRUE),
            First.SLJ.Jump.Height= mean(First.SLJ.Jump.Height, na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s., na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s., na.rm=TRUE),
            RSI.Modified..m.s.= mean(RSI.Modified..m.s., na.rm=TRUE),
            Max.RSI.Modified..m.s.= mean(Max.RSI.Modified..m.s., na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD, na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s., na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s., na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Concentric.Impulse..Ns.= mean(Concentric.Impulse..Ns.,na.rm=TRUE),
            Concentric.Impulse..Left...Ns.= mean(Concentric.Impulse..Left...Ns.,na.rm=TRUE),
            Concentric.Impulse..Right...Ns.= mean(Concentric.Impulse..Right...Ns.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Position= get_mode(Position),
            .groups = 'drop'
              )   #60 obs 25 vars
```

#Forcedecks (VALD) Volleyball Data
```{r}
#1: TEST TYPE=CMJ
CMJ_V <- subset(CMJ_V, select = c("anon_id","Date","Position","Test.Type","Weight..kg.","Countermovement.Depth..cm.","Peak.Power...BM..W.kg.","Landing.RFD..N.s.", "Eccentric.Concentric.Mean.Force.Ratio", "Peak.Landing.Force..Asym...N.", "Peak.Power..W.","Concentric.Maximum.RFD..N.s.","Concentric.Maximum.RFD..Right...N.s.","Concentric.Maximum.RFD..Left...N.s.","Eccentric.Mean.Braking.Force..N.", "Max.Jump.Height..cm.", "Max.Concentric.Peak.Force..N.", "Relative.Concentric.Power.Bilateral",
"Relative.Eccentric.Power.Bilateral","Max.Peak.Power...BM", "Max.Eccentric.Peak.Force..N.",
"Max.RSI.Modified.VALD","Max.Concentric.RPD..N.s.", "Max.Concentric.Peak.Velocity..m.s.", "Max.Eccentric.Peak.Velocity..m.s.", "Max.Peak.Landing.Force..N.", "MAX.Concentric.Peak.Power..W.","Max.CMJ.Jump.Height..cm.",
"Max.CMJ.RSI.Modified..m.s.","Max.CMJ.Concentric.RPD..N.s.","Max.CMJ.Peak.Power..W.","Max.CMJ.Peak.Landing.Force..N.","CMJ.Concentric.Relative.Peak.Power...Difference.from.First","CMJ.Eccentric.Relative.Peak.Power...Difference.from.First","Eccentric.Asymmetry","Concentric.Asymmetry"
))        #5332 obs & 36 vars 

#possible different final dataset 
cmj_v <- CMJ_V %>%
  na.omit()  #457 obs

CMJ_V[CMJ_V==""] <- NA

#combining values from different observations into one
CMJ_v <- CMJ_V %>%
mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  filter(Test.Type=="CMJ") %>%   #add to others even tho might not need
  group_by(anon_id, Date) %>%
  summarise(Weight..kg. = mean(Weight..kg., na.rm = TRUE),
            Countermovement.Depth..cm.= mean(Countermovement.Depth..cm., na.rm=TRUE),
            Peak.Power...BM..W.kg.= mean(Peak.Power...BM..W.kg., na.rm=TRUE),
            Landing.RFD..N.s.= mean(Landing.RFD..N.s.,na.rm=TRUE),
            Eccentric.Concentric.Mean.Force.Ratio= mean(Eccentric.Concentric.Mean.Force.Ratio,na.rm=TRUE),
            Peak.Landing.Force..Asym...N.= mean(Peak.Landing.Force..Asym...N.,na.rm=TRUE),
            Peak.Power..W.= mean(Peak.Power..W.,na.rm=TRUE),
            Concentric.Maximum.RFD..N.s.= mean(Concentric.Maximum.RFD..N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Right...N.s.= mean(Concentric.Maximum.RFD..Right...N.s.,na.rm=TRUE),
            Concentric.Maximum.RFD..Left...N.s.= mean(Concentric.Maximum.RFD..Left...N.s.,na.rm=TRUE),
            Eccentric.Mean.Braking.Force..N.= mean(Eccentric.Mean.Braking.Force..N.,na.rm=TRUE),
            Max.Jump.Height..cm.= mean(Max.Jump.Height..cm.,na.rm=TRUE),
            Max.Concentric.Peak.Force..N.= mean(Max.Concentric.Peak.Force..N.,na.rm=TRUE),
            Relative.Concentric.Power.Bilateral= mean(Relative.Concentric.Power.Bilateral,na.rm=TRUE),
            Relative.Eccentric.Power.Bilateral= mean(Relative.Eccentric.Power.Bilateral,na.rm=TRUE),
            Max.Peak.Power...BM= mean(Max.Peak.Power...BM,na.rm=TRUE),
            Max.Eccentric.Peak.Force..N.= mean(Max.Eccentric.Peak.Force..N.,na.rm=TRUE),
            Max.RSI.Modified.VALD= mean(Max.RSI.Modified.VALD,na.rm=TRUE),
            Max.Concentric.RPD..N.s.= mean(Max.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.Concentric.Peak.Velocity..m.s.= mean(Max.Concentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Eccentric.Peak.Velocity..m.s.= mean(Max.Eccentric.Peak.Velocity..m.s.,na.rm=TRUE),
            Max.Peak.Landing.Force..N.= mean(Max.Peak.Landing.Force..N.,na.rm=TRUE),
            MAX.Concentric.Peak.Power..W.= mean(MAX.Concentric.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Jump.Height..cm.= mean(Max.CMJ.Jump.Height..cm.,na.rm=TRUE),
            Max.CMJ.RSI.Modified..m.s.= mean(Max.CMJ.RSI.Modified..m.s.,na.rm=TRUE),
            Max.CMJ.Concentric.RPD..N.s.= mean(Max.CMJ.Concentric.RPD..N.s.,na.rm=TRUE),
            Max.CMJ.Peak.Power..W.= mean(Max.CMJ.Peak.Power..W.,na.rm=TRUE),
            Max.CMJ.Peak.Landing.Force..N.= mean(Max.CMJ.Peak.Landing.Force..N.,na.rm=TRUE),
            CMJ.Concentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Concentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            CMJ.Eccentric.Relative.Peak.Power...Difference.from.First= mean(CMJ.Eccentric.Relative.Peak.Power...Difference.from.First,na.rm=TRUE),
            Position= get_mode(Position),
            Eccentric.Asymmetry= get_mode(Eccentric.Asymmetry),
            Concentric.Asymmetry= get_mode(Concentric.Asymmetry),
            .groups = 'drop'
            
              )         # 1416 obs & 32 vars -> 1801 obs w/ Position Eccentric + Concentric Asymmetry

CMJ_vb <- CMJ_v %>%
  na.omit()   #508 obs & 32 vars  -> 173 obs w/ Position Eccentric + Concentric Asymmetry
```

```{r}
#WRONG AND SHOULD CHANGE 
#2:TEST TYPE= IMTP
IMTP_V <- Forcedecks_V %>%
  filter(Test.Type=="IMTP") %>% #1241
  select("anon_id", "Date", "Baseline.Force..N.","Baseline.Force..Asym...N.", "Peak.Vertical.Force..N.", "Peak.Vertical.Force...BM..N.kg.", "Peak.Vertical.Force..Asym...N.", "RFD...50ms..N.s.", "RFD...50ms..Asym...N.s.", "RFD...100ms..N.s.", "RFD...100ms..Asym...N.s.", "RFD...150ms..N.s.", "RFD...150ms..Asym...N.s.", "RFD...200ms..N.s.", "RFD...200ms..Asym...N.s.", "Net.Peak.Vertical.Force..N.s.", "Net.Peak.Vertical.Force..Asym...N.s.", "Force.at.200ms...BM..N.kg.", "Eccentric.Asymmetry", "Concentric.Asymmetry","MEAN.Eccentric.Mean.Force.Left", "MEAN.Eccentric.Mean.Force.Right", "MEAN.Concentric.Mean.Force.Left", "MEAN.Concentric.Mean.Force.Right", "MEAN.Concentric.Mean.Force.Asym", "MEAN.Peak.Landing.Force.Asym",  "MEAN.Relative.Peak.Power", "MEAN.CMJ.Concentric.Force", "MEAN.Peak.Landing.Force.Left", "MEAN.Peak.Landing.Force.Right") %>%
  na.omit()   #249 obs & 30 vars

#Absolute.Impulse.50ms..Ns., Absolute.Impulse.50ms..Asym…Ns., Absolute.Impulse.100ms..Ns., Absolute.Impulse.100ms..Asym…Ns., Absolute.Impulse.150ms..Ns., Absolute.Impulse.150ms..Asym…Ns., Absolute.Impulse.200ms..Ns., Absolute.Impulse.200ms..Asym…Ns.
```

```{r}
#3: TEST TYPE= HJ
HJ_V <- Forcedecks_V %>%
  filter(Test.Type=="HJ")  %>%  #367 obs
  select("anon_id", "Date", "Corrected.Standing.Weight..Kg.", "Mean.RSI..Flight.Contact.Time.", "Active.Stiffness..N.m.", "Best.Average.Force..N.", "Best.Average.Force..Asym...N.", "Best.Time.to.Peak.Force..ms.", "Best.Contact.Time..ms.", "Best.Flight.Time..ms.", "Best.Impulse..N.s.", "Best.Peak.Force..N.", "Best.Peak.Force..Asym...N.", "Mean.Average.Force..N.", "Mean.Average.Force..Asym...N.", "Mean.Landing.RFD..N.s.", "Mean.Landing.RFD..Asym...N.s.", "Stiffness.Fatigue....", "Stiffness.Fatigue..Asym.....", "Max.Jump.Height..cm.") %>%
  na.omit()
                 #352 obs & 20 vars
```

Basketball, soccer, lacrosse, volleyball for every single major Test.Type.


4. VALD- Performance Test datasets
```{r}
Performance_B <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WBB.csv")
Performance_S <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WSOC.csv")
Performance_L <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WLAx (1).csv")
Performance_V <- read.csv("~/T2P2-data-analysis-2025/data-sets/VALD - Performance Test WVB.csv")
```

Of all performance datasets, only basketball and volleyball have positional norm information but are also missing a lot of positional information which makes this column sparse. None of the datasets contain rotational data or razor data. Contain small amounts of iso prone data.
```{r}
#Basketball preformance data 
#selects variables of interest, nordic and hip data, and cleans up position variable
performance_B <- subset(Performance_B, select = c('anon_id', 'Date', 'Athlete.Bodyweight..kg.','Sport.Position', 'Side', 'Repetitions', 'Maximum.Force', 'Average.Force', 'Impulse', 'Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean', 'Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Position = case_when(
    Sport.Position == '' ~ 'None',
    Sport.Position == '|Guard' ~ 'Guard',
    str_detect(Sport.Position, "Primary:") ~ 'Forward',
    TRUE ~ Sport.Position
  )) %>%
  select(-Sport.Position)

#makes the date column a 'date' object for later and omits NA values for nordic dataset
performance_B_nordic <- performance_B %>%
  select(-c('Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()

#makes the date column a 'date' object for later and omits NA values for hip dataset
performance_B_hip <- performance_B %>%
  select(-c('Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()
```
Need to either get more data on positions of players or get rid of percent difference from norm since it has a lot of NAs. Could consider athlete bodyweight as well to put strength relatively but also has missing data and has negative values? Also missing a decent amount of z-score observations. 147 observations with percent difference norm and z-scores. 550 with only z-scores. 592 with neither percent difference norm or z-scores. If we include athlete bodyweight and exclude percent difference norm and z-scores there are 521 observations. 

Hip dataset contains 109 observations including the bodyweight variable, 129 otherwise. 

```{r}
#Soccer preformance data
#selects variables of interest, nordic and hip data, and cleans up position variable
performance_S <- subset(Performance_S, select = c('anon_id', 'Date', 'Athlete.Bodyweight..kg.','Sport.Position', 'Side', 'Repetitions', 'Maximum.Force', 'Average.Force', 'Impulse', 'Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean', 'Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Position = case_when(
    Sport.Position == '' ~ 'None',
    TRUE ~ Sport.Position
  )) %>%
  select(-Sport.Position)

#makes the date column a 'date' object for later and omits NA values for nordic dataset
performance_S_nordic <- performance_S %>%
  select(-c('Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()

#makes the date column a 'date' object for later and omits NA values for hip dataset
performance_S_hip <- performance_S %>%
  select(-c('Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()
```
Does not contain any information about positional norms so there are no percent differences from norm. 421 observations with z-scores. 483 observations without z-scores. There are 400 observations if bodyweight is included.  

Hip dataset contains only 39 observations with body weight included and without it included. 

```{r}
#Lacrosse preformance data
#selects variables of interest, nordic and hip data, and cleans up position variable
performance_L <- subset(Performance_L, select = c('anon_id', 'Date', 'Athlete.Bodyweight..kg.','Sport.Position', 'Side', 'Repetitions', 'Maximum.Force', 'Average.Force', 'Impulse', 'Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean', 'Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Position = case_when(
    Sport.Position == '' ~ 'None',
    TRUE ~ Sport.Position
  )) %>%
  select(-Sport.Position)

#makes the date column a 'date' object for later and omits NA values for nordic dataset
performance_L_nordic <- performance_L %>%
  select(-c('Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()

#makes the date column a 'date' object for later and omits NA values for hip dataset
performance_L_hip <- performance_L %>%
  select(-c('Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()
```

Also does not contain any information about positional norms so there are no percent differences from norm. 855 observations with z-scores. 929 observations without z-scores. 810 observations if bodyweight is included.  

Hip dataset contains 100 observations with bodyweight included and excluded. 

```{r}
#4Volleyball preformance data
#selects variables of interest, nordic and hip data
performance_V <- subset(Performance_V, select = c('anon_id', 'Date', 'Athlete.Bodyweight..kg.','Sport.Position', 'Side', 'Repetitions', 'Maximum.Force', 'Average.Force', 'Impulse', 'Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean', 'Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Position = case_when(
    Sport.Position == '' ~ 'None',
    TRUE ~ Sport.Position
  )) %>%
  select(-Sport.Position)

#makes the date column a 'date' object for later and omits NA values for nordic dataset
performance_V_nordic <- performance_V %>%
  select(-c('Hip.Abd.Left.MEAN', 'Hip.Abd.Right.MEAN', 'Hip.Abduction.MEAN.Imbalance', 'Hip.Abd.Left.MAX', 'Hip.Abd.Right.MAX', 'Hip.Abduction.MAX.Imbalance', 'Hip.Add.Left.MEAN', 'Hip.Add.Right.MEAN', 'Hip.Adduction.MEAN.Imbalance', 'Hip.Add.Left.MAX', 'Hip.Add.Right.MAX', 'Hip.Adduction.MAX.Imbalance', 'Bilateral.Hip.Abduction.Adduction.Ratio')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()

#makes the date column a 'date' object for later and omits NA values for hip dataset
performance_V_hip <- performance_V %>%
  select(-c('Nordic.Left.MEAN', 'Nordic.Right.MEAN', 'Nordic.MEAN.Imbalance', 'Nordic.Left.MAX', 'Nordic.Right.MAX', 'Nordic.MAX.Imbalance', 'Trend', 'Nordic...Difference.from.First.Test', 'Maximum.Nordic.Bilateral.Mean', 'Relative.Maximum.Nordic.Bilateral.Mean')) %>%
  mutate(Date = as.Date(Date, format = '%m/%d/%Y')) %>%
  na.omit()
```

There are a lot of observations with the player playing multiple positions, need to figure out how to filter down when analyzing data by positions. 

Does have positional norm information but a lot of missing values. Only 33 observations with percent difference norm and z-scores. 356 observations with only z-scores. 393 observations without either percent difference norm or z-scores. 325 observations if bodyweight is included.

Hip data contains 180 observations with bodyweight included, 220 otherwise.


## Exploratory Analysis 


#Q1: What are meaningful thresholds for strength metrics as they relate to lower body injury risk? 
a. For example, one of our strength coaches uses 5 times body weight (kg) in Newtons for hamstring strength as a cutoff for injury risk. Is this supported by our data? 

-looking for meaningful strength metrics to lower body injury risk
-create thresholds fir these metrics by each position in each sport

#Q2:What trends in strength do we see across the women’s team sports? 
a. Soccer, lacrosse, volleyball, basketball. 


